---
title: "Homework 6"
author: "Gauri Bhatkhande"
date: "09/12/2020"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
```

## Problem 2

**Importing the dataset**
```{r}
birthweight_data = read_csv(file = "./data/birthweight.csv")
```

**Cleaning the dataset** 

```{r}
birthweight_data %>%
  sapply(function(x) sum(is.na(x))) 

birthweight_df = 
  birthweight_data %>%
  mutate(
    babysex = as.factor(babysex) 
  )



```


**Proposed regression model for the outcome birthweight.** 

Based on hypothesis found in literature, I have selected the following variables as predictors: 
Mother's weight gain during pregnancy (pounds) (wtgain)
Gestational age in weeks (gaweeks)
Smoking status - Average number of cigarettes smoked per day during pregnancy (smoken)
Baby’s sex (babysex)

This model was decided based on literature evidence that suggested an association between these predictors and the baby's birthweight. 

**Building the proposed model:** 

```{r}
lin_model = lm(bwt ~ wtgain + gaweeks + smoken + babysex, data = birthweight_df)

lin_model %>%
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  mutate(term = str_replace(term, "babysex2", "baby's sex: female")) %>% 
  knitr::kable(digits = 3)
```




